---
title: "Strawberries :("
author: "Amie Thomas"
date: "2023-10-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("svglite")
library(knitr)  
library(kableExtra)
library(tidyverse)
library(stringr)
```


<Loading data>
```{r}
strawberry <- read.csv("strawberry(1).csv")

#looking at data before
glimpse(strawberry)

```

< Get rid of single value columns. Haviland's code >
```{r}
drop_single_val <- function(df){
  col_val <- NULL
  col_name <- NULL
suppressWarnings({
  for(i in 1:dim(df)[2]){
if((df |> distinct(df[,i]) |> count()) == 1){
  col_name = c(col_name, colnames(df[i]))
  col_val = c(col_val, df[1,i])  
} }
})

if(is.null(col_name)){return("No Columns to drop")}else{
   col_val = unlist(col_val)
   attributes(col_val) = NULL
   drp = data.frame(col_name, col_val)
   return(drp)
   }
}

str <- drop_single_val(strawberry)

str <- str$col_name

strawberry <- strawberry |> select(!all_of(str))
```

```{r}
#take a gander at the data after
glimpse(strawberry)
```

```{r}
#check if every row has is associated with a state
###question to self: why do we need to do this
state_all <- strawberry |> group_by(State) |> count()

### test if every row is associated with a state by summing the 
## counts and testing for equality with the total rows in the 
## data frame

if(sum(state_all$n) == dim(strawberry)[1]){print("Every row has value in the State column.")}
```

So far, Haviland has removed missing values from the columns. And I think checked for missing values in the rows. Why did Haviland find the state with the most rows. Ask about the difference in his notations. Is this a common way of notation

```{r}
#which state has the most rows? 
#how does this information help us?
state_max <- state_all$State[which(state_all$n ==  max(state_all$n)  )]


```

#Havilands notes
<!-- ## Separate CENSUS and SURVEY into two Data Frames -->

<!-- In the strawberry data frame,  -->
<!-- The CENSUS rows contains marketing, sales, and productiong data.  The SURVEY rows contain rows which may be redundant with the CENSUS rows and chemical application rows. -->

<!-- After splitting CENSUS and SURVEY rows into two data frames,  -->
<!-- finish organizing the CENSUS data first.  Then, organize the -->
<!-- SURVEY data frame splitting the marketing, and production data from the chemical application data. -->


```{r}
strwb_survey <- strawberry |> filter(Program == "SURVEY")
strwb_census <- strawberry |> filter(Program == "CENSUS")
```

Need to separate the columns that have more than one type of information. Domain item and Domain category
<!-- Column separators in CENSUS: ",", "-", ":" -->

<!-- ### Separate `Data Item` into columns by "," -->

#separate_wider_delim i think separates strings in to columns by delimiters 

```{r}
strwb_census <- strwb_census |>
  separate_wider_delim(  cols = `Data Item`,
                         delim = ",",
                         names = c("Fruit",
                                 "temp1",
                                 "temp2",
                                 "temp3"),
                         too_many = "error",
                         too_few = "align_start"
                       )
#what does the too many and too few do in this

strwb_census <- strwb_census |>
  separate_wider_delim(  cols = temp1,
                         delim = " - ",
                         names = c("crop_type",
                                 "prop_acct"),
                         too_many = "error",
                         too_few = "align_start"
                       )

## trim the strings
## you can see which columns contain string values that need
## to have leading or trailing spaces that need to be trimmed.


# glimpse(strwb_census)

strwb_census$crop_type <- str_trim(strwb_census$crop_type, side = "both")

strwb_census$temp2 <- str_trim(strwb_census$temp2, side = "both")

strwb_census$temp3 <- str_trim(strwb_census$temp3, side = "both")
```

< Fresh market column >

```{r}
#| label: create a fresh market column
#| echo: false
#| eval: true

## make a copy of the temp2 column named `Fresh Market`.
strwb_census <- strwb_census |> mutate(`Fresh Market` = temp2, .after = temp2)

## Remove cells in `Fresh Market` column 
##   that begin "MEASURED"
strwb_census$`Fresh Market` <- strwb_census$`Fresh Market` |> str_replace( "^MEA.*", "")

## Remove cells in `Fresh Market` column 
##   that begin "PROCESSING" 
strwb_census$`Fresh Market` <- strwb_census$`Fresh Market` |> str_replace( "^P.*", "")

## substitute a space for NA in `Fresh Market` column
strwb_census$`Fresh Market`[is.na(strwb_census$`Fresh Market`)] <- ""  

## in temp2 column, remove cells that begin "FRESH"
 strwb_census$temp2 <- strwb_census$temp2 |> str_replace("^F.*", "")

## Now fix the entries in the `Fresh Market` column
##   Remove "FRESH MARKET - " from the cells
strwb_census$`Fresh Market` <- strwb_census$`Fresh Market` |> str_replace("^FRESH MARKET - ", "")

```

<!-- ## Create a "Process Market" column -->

```{r}
#| label: make process market column
#| echo: false

## Make a copy of temp2 named `Process Market`
strwb_census <- strwb_census |> mutate(`Process Market` = temp2, .after = temp2)

## remove `Process Market` cells beginning "MEASURED"
strwb_census$`Process Market` <-  strwb_census$`Process Market` |> str_replace("^MEA.*", "")

## substitute space for NA in `Process Market` column
strwb_census$`Process Market`[is.na(strwb_census$`Process Market`)] <- ""

## In temp2, remove cells that begin "PROCESSING"
strwb_census$temp2 <- strwb_census$temp2 |> str_replace("^P.*", "")

## In `Processing Market`, remove "PROCESSING - " from cells
strwb_census$`Process Market` <-  strwb_census$`Process Market` |> str_replace("PROCESSING - ", "") 
```

<!-- Remove NA's from prop_acct, temp2, and temp3 -->

```{r}
## substitute a space for NA in prop_acct column
strwb_census$prop_acct[is.na(strwb_census$prop_acct)] <- "" 

## substitute a space for NA in temp2 column
strwb_census$temp2[is.na(strwb_census$temp2)] <- "" 

## substitute a space for NA in temp2 column
strwb_census$temp3[is.na(strwb_census$temp3)] <- "" 

```

<!-- Combine temp2 with temp3 to create Metric column -->
<!-- remove  -->
<!-- relocate columns -->

```{r}
strwb_census <- strwb_census |> unite(temp2, temp3, col="Metric", sep="")

## Now fix the entries in the Metric column
##   Remove "MEASURED IN " from the cells
strwb_census$Metric <- strwb_census$Metric |> str_replace("MEASURED IN ", "")

## move Metric to the end
strwb_census <- strwb_census |> relocate(Metric, .before = Domain)

strwb_census <- strwb_census |> relocate(`Process Market`, .before = Metric)

strwb_census <- strwb_census |> rename(Totals = prop_acct)

#drop_one_value_col(strwb_census)

```


```{r}

```



